---
title: "DPH101 Biostatistics I: Week 03 R Workshop"
author: "Elmer V Villanueva"
date: "September 28, 2018"
output: html_document
---

---
# Comment
# setwd("F:/Dropbox/00 - Working Folder/Teaching/DPH101/2018-2019/Week 03/R03 R Workshop")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic Probability
## Binomial Probability
$$ P(X=k) = \binom{n}{k} p^k (1-p)^{(n-k)} $$

Given $p = 0.25$:

1. What is the probability that you observe exactly 2 successes in 12 trials?
```{r}
dbinom(x = 2, size = 12, prob = 0.25)
```

2. What is the probability of observing two successes at most?
```{r}
dbinom(x = 0, size = 12, prob = 0.25) + dbinom(x = 1, size = 12, prob = 0.25) + dbinom(x = 2, size = 12, prob = 0.25)
pbinom(q = 2, size = 12, prob = 0.25, lower.tail = TRUE)
```
*pbinom is accumulation, using "q" lower.tail means give me the lower tail that is smaller than 2.

3. What is the probability of observing 3 or more successes?
```{r}
1 - pbinom(q = 2, size = 12, prob = 0.25, lower.tail = TRUE)
pbinom(q = 2, size = 12, prob = 0.25, lower.tail = FALSE)
```

## Poisson probabilities
You should review the `dpois()` and `ppois()` functions.

## Normal probabilities
You should review the `dnorm()` and `pnorm()` functions.
*above two is the HOMEWORK.

# Higher graphics
## Preparing for the session
### Set your working directory
```{r echo=TRUE}
setwd("~/R_Directory/DPH101_W3")
```

### Download the data file from ICE onto your working directory
For this session, we will be using the `GLOW500FINAL.RData` file.

### Load the data file onto the global environment
```{r echo=TRUE}
load("GLOW500FINAL.RData")
```

The dataframe is called `GLOW500_WORK`. Confirm that the data have been loaded and are reasonably correct.
```{r echo=TRUE}
str(GLOW500_WORK)
head(GLOW500_WORK)
attach(GLOW500_WORK)
```
*str means give me the structure of the data.
*notice the first column after variale name is int(whole number), num(can contain decimals), for factor(no more a integer, change the integer into some practical meanings, 1 or 2 means others things).
*LEARN how to change numeric into factor.

### Install and load the ggplot2 package
```{r eval=TRUE}
if (!require("ggplot2")) install.packages("gglot2", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library(ggplot2)

if (!require("ggthemes")) install.packages("ggthemes", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library(ggthemes)

if (!require("scales")) install.packages("scales", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library(scales)

if (!require("showtext")) install.packages("showtext", repos = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/")
library(showtext)
```

## Histograms
The original histogram we produced in class was this.
```{r echo=TRUE}
hist(HEIGHT, breaks=15, col="tan", xlab="Height (cm)",
     main="Histogram of Heights in GLOW500")
```

We can produce one in `ggplot2` easily.

1. Initialise a plot by telling `ggplot2` that `GLOW500_WORK` is our data frame, and specify the variables on each axis. We then render this as histogram by adding the `geom_histogram` command. Note, specifically, that the for of the y axis is specified in this command.

```{r echo=TRUE}
hist1 <- ggplot(aes(x = HEIGHT),
                data = GLOW500_WORK) + 
         geom_histogram(aes(y = ..count..),
                        bins = 15,
                        colour = "black", 
                        fill = "tan")
hist1
```

2. Let us add titles.
```{r echo=TRUE}
hist1 <- hist1 + scale_x_continuous(name = "Height (cm)") +
         scale_y_continuous(name = "Frequency") +
         labs(title = "Frequency histogram of height",
              subtitle = "Source: Global Longitudinal Study of Osteoporosis in Women")
hist1
```

3. Instead of histograms, we can ask `ggplot2` to produce density plots by replacing `geom_histogram()` with `geom_density()`
```{r echo=TRUE}
hist1 <- ggplot(aes(x = HEIGHT),
                data = GLOW500_WORK) + 
         geom_density(colour = "black", 
                      fill = "tan") + 
         scale_x_continuous(name = "Height (cm)") +
         scale_y_continuous(name = "Density") +
         labs(title = "Density plot of height",
              subtitle = "Source: Global Longitudinal Study of Osteoporosis in Women")
hist1
```

4. We can produce histograms and densities by different factors. For example, let's produce densities of HEIGHT across levels of BMICAT.
```{r echo=TRUE}
hist1 <- ggplot(aes(x = HEIGHT, fill = BMICAT),
                data = GLOW500_WORK) + 
         geom_density(colour = "black", 
                      fill = "tan") + 
         scale_x_continuous(name = "Height (cm)") +
         scale_y_continuous(name = "Density") +
         labs(title = "Density plot of height by BMI",
              subtitle = "Source: Global Longitudinal Study of Osteoporosis in Women") +
         facet_grid(. ~ BMICAT)
hist1
```

5. Finally, we can superimpose the densities on each other.
```{r echo=TRUE}
hist1 <- ggplot(aes(x = HEIGHT, fill = BMICAT),
                data = GLOW500_WORK) + 
         geom_density(position = "identity",
                      alpha = 0.3) + 
         scale_x_continuous(name = "Height (cm)") +
         scale_y_continuous(name = "Density") +
         labs(title = "Density plot of height by BMI",
              subtitle = "Source: Global Longitudinal Study of Osteoporosis in Women")
hist1
```


6. This might be a little hard to read, so we can adjust the `fill` option to `NA`, change the line colours and make it thicker.
```{r echo=TRUE}
hist1 <- ggplot(aes(x = HEIGHT, colour = BMICAT),
                data = GLOW500_WORK) + 
         geom_density(position = "identity",
                      fill = NA,
                      size = 1) + 
         scale_x_continuous(name = "Height (cm)") +
         scale_y_continuous(name = "Density") +
         labs(title = "Density plot of height by BMI",
              subtitle = "Source: Global Longitudinal Study of Osteoporosis in Women")
hist1
```

I'm not quite sure that this is better, though.

## Bar Charts
In class, we learned how to produce a bar chart of BMICAT by AGECAT using the `graphics` package.
```{r echo=TRUE}
table(GLOW500_WORK$BMICAT, GLOW500_WORK$AGECAT)
AGECAT.2 <- table(GLOW500_WORK$BMICAT, GLOW500_WORK$AGECAT)
barplot(AGECAT.2,
        xlab="Age Group (years)",
        ylab="Frequency",
        main="Distribution of BMI by Age",
        legend.text=TRUE)
```

Let's begin building the chart in `ggplot2`. 

1. Convert `AGECAT.2` from a table to a data frame.
```{r echo=TRUE}
AGECAT.2
class(AGECAT.2)
AGECAT.2.DF <- as.data.frame(AGECAT.2) 
AGECAT.2.DF
class(AGECAT.2.DF)
```

2. Initialise a plot by telling `ggplot2` that `AGECAT.2.DF` is our data frame, and specify the variables on each axis. We then render this as an bar plot by adding the `geom_area` command.
```{r echo=TRUE}
bar1 <- ggplot(aes(y = Freq, 
                   x = AGECAT.2.DF[,2], fill = AGECAT.2.DF[,1]), 
               data = AGECAT.2.DF) + 
        geom_col()
bar1
```

3. Change the position of the legend. 
To label the bars, we add the `label` argument.
```{r echo=TRUE}
bar2 <- bar1 + theme(legend.position = "bottom", 
                     legend.direction = "horizontal",
                     legend.title = element_blank())
bar2
```

4. Let us adjust the axis titles and produce a title for the bar chart.
```{r echo=TRUE, warning=FALSE}
bar3 <- bar2 + labs(title = "Distribution of BMI by AGE",
                    subtitle = "Source: Global Longitudinal Study of Osteoporosis in Women") + 
        labs(x = "Age Group (years)", y = "Frequency")
bar3
```

5. We can use themes to make the chart appear similar to those, say, produced by *The Economist*.
```{r echo=TRUE, warning=FALSE}
library(ggthemes)
bar4 <- bar3 + labs(title = "Distribution of BMI by AGE",
                    subtitle = "Source: Global Longitudinal Study of Osteoporosis in Women") + 
        labs(x = "Age Group (years)", y = "Frequency") +
        theme_economist() + 
        scale_color_economist() +
        theme(axis.line.x = element_line(size = .5, 
                                         colour = "black"),
              legend.position = "bottom",
              legend.direction = "horizontal",
              legend.title = element_blank(),
              plot.title = element_text(family = "Roboto Condensed"),
              text = element_text(family = "Roboto Condensed")) +
        guides(fill = guide_legend(reverse = T))
bar4
```


